\mytitle{}

### 2019-09-25 paper/paper-revised.tex
Almost there...

\cbstart
\subsection{Contributions}
This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
These features have been suggested by past studies into DDoS prevention, and have been independently tested in their own contexts.
Our study is the first attempt to quantify the individual efficacy of each in an RL-focussed setting.

We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
Observing that past work has been compared only against AIMD \cite{DBLP:journals/ton/YauLLY05}, we additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS defences.
\cbend

@@

\begin{figure}
	\centering
	\resizebox{0.5\linewidth}{!}{
		\begin{tikzpicture}		
		% Top half
		\def\startacs{
			{0.8},
			{0.3},
			{-0.4}%
		}

@@

\subsection{Threat Model}
An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to a host which connects to it, and they are expected to act rationally in its pursuit.
Threat actors are external and act intentionally, aren't expected to be 
\emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.

@@

\subsection{System Design and Assumptions}
A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model or agent design, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.
\cbend

@@

\cbstart
Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
Only the ingress/egress nodes from a network need to be OpenFlow-enabled, because it is advantageous to perform filtering as close to a flow's source as possible.
In a traditional network, each agent has exclusive control over its switch's tables.
In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
The main difference is that a traditional network needs this additional hardware, and does not allow for an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
\cbend

@@

Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.

@@

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.

@@

\begin{figure}
	\centering
	\vspace{-0.75cm}
	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
	\vspace{-0.75cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
		\label{fig:tcp-laf-feature-plots}
	}
\end{figure}

@@

Hosts send RTP traffic with Salsa20 encrypted payloads---\SI{20}{\milli\second} audio frames at \SI{96}{\kilo\bit\per\second}.
We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers, and estimate the size of missed packets by taking an exponentially-weighted moving average over known sizes.
Hosts punctuate audio frames with a 4-byte keepalive every \SI{5}{\second}.
All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
There is no peer-to-peer traffic---the server acts as a TURN relay for all hosts.
%?? Reflective factor among \emph{authenticated hosts}.

We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
To match the target upload rate assigned to each host, it runs enough individual sessions to meet the target data rate.

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
\cbstart
\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.
\cbend

@@

We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
\cbstart
Marl is already known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} approach to DDoS mitigation, although the state of the art has long since moved on.
To paint a more current picture, we compare our work against an effective modern-day approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
We compare ourselves against this approach to position our new RL agent designs against the modern state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \SI{5}{\second}.
\cbend

@@

\subsection{Single Destination}\label{sec:single-dest}
%?? Move description of tree topol to here.
The network is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
We then have $N_{\mathit{hosts}} = k \ell m n$.
\Cref{fig:marl-topol} demonstrates this.
%Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a}, it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
%It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
%We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
We configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
This is a slight simplification of \citeauthor{DBLP:journals/eaai/MalialisK15}'s \textquote{online} experiment \cite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
%The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
%Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
%$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
%The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
%Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
%These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

\begin{figure}
	\centering
	\resizebox{0.8\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=uofgsandstone},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node[treenode, label={[texts]above:Server}] (root) {}
		child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
			child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
				child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
					child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
						child [treeline] { node [treenode, external] (extern) {}
							child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
							child [hideline] { node [hidenode, externalhide] (endhost) {} }
						}
					}
					child [hideline] { node [hidenode, loadhide] (endagent) {} }
				}
				child [hideline] { node [hidenode] (endinter) {} }
			}
			child [hideline] { node [hidenode] (endteaml) {} }
			edge from parent
			node[below, labeltexts] {$U_s$}
		};

		%\draw[-] (teaml) -- (endteaml);
		\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of kdots] {$k$};
		\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
		\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of mdots] {$m$};
		\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of ndots] {$n$};
		\end{tikzpicture}
	}
	\caption{
		Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
		%	Empty nodes are considered to be internal.
		Red nodes are external, and each blue node hosts an agent.
		\label{fig:marl-topol}
	}
\end{figure}

\subsection{Multiple Destinations}
The previous topology allows for direct comparison against the state-of-the-art, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
In contrast, the fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.

%?? Come up with description of fat-tree (multi-dest) topol.
%?? Why fat tree? regularly appears in modern datacentres.
%?? $k=4$ fat-tree , with one pod hosting two servers $s_0,s_1$.
We use a $k=4$ fat-tree, with one pod hosting two servers $s_0$ and $s_1$.
$n$ external hosts connect through each core switch (where agents are hosted), and communicate with $s_0, s_1$ uniformly randomly.
Both servers host identical services.
We set $n \in \{6, 12, 24, 48\}$ hosts per learner (keeping $N_{\mathit{hosts}}$ identical to each tier of the single-host topology), and restrict $U_{s_0} = U_{s_1} = U_s / 2$.

@@

$U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
All links had unbounded capacity, save for server-switch connections.
These parameter choices match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

@@

We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold and underlining the best RL-based model.
Average rewards take into account any portions of time that an agent allows illegal system states.
Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.

@@

\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
\cbstart Here, SPIFFY greatly outperforms all of the RL-based methods. \cbend
%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
%?? Discuss multi-dest topol once results available.
We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
\cbstart SPIFFY continues to outperform the RL-based methods. \cbend

>>>>

\subsection{Contributions}
\cbstart
This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
These features have been suggested by past studies, and independently tested in their own contexts.
Our study is the first attempt to quantify the individual efficacy of each in an RL setting.

We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.

@@

\begin{figure}
	\centering
	\resizebox{0.45\linewidth}{!}{
		\begin{tikzpicture}		
		% Top half
		\def\startacs{
			{0.8},
			{0.3},
			{-0.4}%
		}

@@

\subsection{Threat Model}
An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to hosts, and they are expected to act rationally in its pursuit.
Threat actors are external and act intentionally, aren't expected to be 
\emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
We assume that attacks will be volumetric DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.

@@

\subsection{System Design and Assumptions}
A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, \cbend $U_s$.

@@

\cbstart
Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
In a traditional network, each agent has exclusive control over its switch's tables.
In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
\cbend

@@

Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we expect that the simplicity of tile-based policy computation will also greatly aid interpretability of anomalous action choices.

@@

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
\cbstart
Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
\cbend

@@

\begin{figure}
	\centering
%	\vspace{-0.25cm}
	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
	\vspace{-0.75cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
		\label{fig:tcp-laf-feature-plots}
	}
\vspace{-0.6cm}
\end{figure}

@@

Hosts send RTP traffic with Salsa20 encrypted payloads---\SI{20}{\milli\second} audio frames at \SI{96}{\kilo\bit\per\second}.
We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers, and estimate the size of missed packets by taking an exponentially-weighted moving average over known sizes.
Hosts punctuate audio frames with a 4-byte keepalive every \SI{5}{\second}.
All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
There is no peer-to-peer traffic---the server acts as a TURN relay for all hosts.
%?? Reflective factor among \emph{authenticated hosts}.
We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
To match the target upload rate assigned to each host, it runs enough individual sessions to meet the target data rate.

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
\cbstart
Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.
\cbend

@@

We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
\cbstart
Marl is already known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} approach to DDoS mitigation, although the state of the art has long since moved on.
To paint a more current picture, we compare our work against an effective modern-day approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
We compare ourselves against this approach to position our new RL agent designs against the modern state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.

@@

\subsection{Single Destination}\label{sec:single-dest}
%?? Move description of tree topol to here.
The network is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
We then have $N_{\mathit{hosts}} = k \ell m n$.
\Cref{fig:marl-topol} demonstrates this.
%Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a}, it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
%It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
%We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
We configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
This is a slight simplification of \citeauthor{DBLP:journals/eaai/MalialisK15}'s \textquote{online} experiment \cite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
%The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
%Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
%$U_s$  $k \ell mn+2$ \si{\mega\bit\per\second}.
%The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
%Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
%These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

\begin{figure}
	\centering
	\resizebox{0.7\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=uofgsandstone},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node[treenode, label={[texts]above:Server}] (root) {}
		child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
			child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
				child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
					child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
						child [treeline] { node [treenode, external] (extern) {}
							child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
							child [hideline] { node [hidenode, externalhide] (endhost) {} }
						}
					}
					child [hideline] { node [hidenode, loadhide] (endagent) {} }
				}
				child [hideline] { node [hidenode] (endinter) {} }
			}
			child [hideline] { node [hidenode] (endteaml) {} }
			edge from parent
			node[below, labeltexts] {$U_s$}
		};

		%\draw[-] (teaml) -- (endteaml);
		\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of kdots] {$k$};
		\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
		\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of mdots] {$m$};
		\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
		\node [labeltexts, right = -0.1cm of ndots] {$n$};
		\end{tikzpicture}
	}
	\caption{
		Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
		%	Empty nodes are considered to be internal.
		Red nodes are external, and each blue node hosts an agent.
		\label{fig:marl-topol}
	}
\vspace{-0.75cm}
\end{figure}

\subsection{Multiple Destinations}
The previous topology allows for direct comparison against the state-of-the-art, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
In contrast, the fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
%?? Come up with description of fat-tree (multi-dest) topol.
%?? Why fat tree? regularly appears in modern datacentres.
%?? $k=4$ fat-tree , with one pod hosting two servers $s_0,s_1$.
We use a $k=4$ fat-tree, with one pod hosting two servers $s_0$ and $s_1$.
$n$ external hosts connect through each core switch (where agents are hosted), and communicate with $s_0, s_1$ uniformly randomly.
Both servers host identical services.
We set $n \in \{6, 12, 24, 48\}$ hosts per learner (keeping $N_{\mathit{hosts}}$ identical to each tier of the single-host topology), and restrict $U_{s_0} = U_{s_1} = U_s / 2$.

@@

$U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
All links had unbounded capacity, save for each server-switch.
These parameters match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

@@

We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model underlined.
Average rewards take into account any portions of time that an agent allows illegal system states.
Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.

@@

\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
%?? Discuss multi-dest topol once results available.
We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
In both cases, SPIFFY greatly outperforms the RL-based methods.
